---
title: "Report of Berry Project"
author: "Haoqi Wang"
date: "10/20/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(ggplot2)
library(citation)

opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```
## 1.Overview

In this assignment, I use the data that were collected from the USDA data base, and I choose strawberry as my data set to do the data cleaning, data organization, exploratory data analysis. Finally, I find some correlations between variables. 

## 2.Acquire and read the data

These data were collected from the USDA database selector: <a href="https://quickstats.nass.usda.gov">https://quickstats.nass.usda.gov</a>"

The data were <a href="https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1">stored online</a> and then downloaded as a CSV file.


```{r}

## read the data

ag_data <- read.csv("berries.csv")

## look at number of unique values in each column
ag_data %>% summarize_all(n_distinct) -> aa

## make a list of the columns with only one unique value
bb <- which(aa[1,]==1)

## list the 1-unique value column names 
cn <- colnames(ag_data)[bb]

```


## 3. Choose Strawberries-Data clean and Data organize 

Data selected from the NASS database often has columns without any data or with a single repeated Values.  The berries data had only 8 out of 21 columns containing meaningful data.

This table contains informaton about berries: blueberries, raspberries, and strawberries.

When the data have been cleaned and organized, the three kinds of berries will be separted into tables with the same stucture so that they can be compared.  So, working with Strawberries along demonstrates how the data will be cleaned and organized.
```{r}
## remove the 1-unique columns from the dataset
ag_data %<>% select(-all_of(bb))

aa %<>% select(-all_of(bb)) 


## State name and the State ANSI code are (sort of) redundant
## Just keep the name
ag_data %<>% select(-4)
aa %<>% select(-4) 
kable(head(ag_data)) %>%kable_styling(font_size=12)

#clean the column of Data.Item
sberry <- ag_data %>% filter((Commodity=="STRAWBERRIES") & (Period=="YEAR"))
sberry %<>% select(-c(Period, Commodity))   
    
# Does every Data Item begin with "
sum(str_detect(sberry$`Data.Item`, "^STRAWBERRIES,")) == length(sberry$`Data.Item`)

#Revise the rows in 'Data.Item' 
for(i in 1:length(sberry$Year)){
      sberry$`Data.Item`[i]<-str_replace(sberry$`Data.Item`[i],"STRAWBERRIES - ","STRAWBERRIES, TAME  ")
}
# Does every Data Item begin with "
sum(str_detect(sberry$`Data.Item`, "^STRAWBERRIES,")) == length(sberry$`Data.Item`)


# separate the items in Data.Item
sberry %<>% separate(`Data.Item`, c("S","Type", "meas", "what"), sep = ",") 
sberry %<>% select(-S)
ty <- str_split(sberry$Type, " ", simplify=TRUE)
sberry %<>% separate(Type,c("s1", "Type", "s2", "lab1", "lab2"), " ")

sberry %<>% select(-c(s1,s2)) 
sberry[is.na(sberry)] <- " "  

## clean the column of Domain

sberry %<>% separate(Domain, c("D_left", "D_right"), sep = ", ")

sberry[is.na(sberry)] <- " "

## clean the column of Domain Category
#Revise the rows in 'Data.Item' 
for(i in 1:length(sberry$Year)){
      sberry$`Domain.Category`[i]<-str_replace(sberry$`Domain.Category`[i],"NOT SPECIFIED","NOT SPECIFIED, ")
}
sberry %<>% separate(`Domain.Category`, c("DC_left", "DC_right"), sep = ", ")

## work on DC_left first

sberry %<>% separate(DC_left, c("DC_left_l", "DC_left_r"), sep = ": ")

## now work on DC_right

sberry %<>% separate(DC_right, c("DC_right_l", "DC_right_r"), sep = ": ") 
## remove NA
sberry[is.na(sberry)] <- " "

## remove column DC_left_l

sberry %<>%  select(-DC_left_l) 

## remove column DC_right_l

sberry %<>% select(-DC_right_l)

##check the lab1 lab2
sberry %<>% mutate(label = paste(lab1,lab2)) 

## test for necisity of "chemical" in col D_left

paste(sberry$D_left, sberry$D_right) %>% unique()
# [1] "TOTAL  "              "CHEMICAL FUNGICIDE"   "CHEMICAL HERBICIDE"   "CHEMICAL INSECTICIDE"
# [5] "CHEMICAL OTHER"       "FERTILIZER  "   


## remove "Chemical" and joint the columns

sberry %<>% mutate(D_left = "CHEMICAL", D_left = "") 
sberry %<>% mutate(Chemical=paste(D_left, D_right)) 
sberry %<>% select(-c(D_left, D_right)) 
sberry %<>% select(Year, State, Type, what, meas, label, DC_left_r, DC_right_r, Chemical, Value )

###  Now the problem is that we have entries in both the "what" and "meas" columns
##  that begin  "MEASURED IN"
f1 <- function(a,b){
  if(a){
    return(b)
  }else{
      return("")
  }
}
index_meas <- str_detect(sberry$meas, "MEASURED IN")
sberry %<>% mutate(m_in_1 = unlist(map2(index_meas, sberry$meas, f1))) 
sberry %<>% mutate(meas = str_replace(sberry$meas, "MEASURED IN.*$", ""))

index_what <- str_detect(sberry$what, "MEASURED IN")
sberry %<>% mutate(m_in_2 = unlist(map2(index_what, sberry$what, f1))) 
sberry %<>% mutate(what = str_replace(sberry$what, "MEASURED IN.*$", ""))


### Check for overlaps

sberry %<>% mutate(units = str_trim(paste(m_in_1, m_in_2))) 

sberry$units %>% unique()


## now let's clean it up 

sberry %<>% rename(Avg = what) 
sberry %<>% rename(Marketing = meas, Harvest = label, Chem_family = DC_left_r, Materials = DC_right_r, Measures = units)

colnames(sberry)

sberry %<>% select(Year, State, Type, Marketing, 
                   Measures, Avg, Harvest, Chem_family,
                   Materials, Chemical, Value )


###  these belong in one column

sberry %<>% mutate(Production = str_trim(paste(Marketing, Harvest)))

sberry %<>% select(Year, State, Type, Production, Measures,
                   Avg, Chem_family, Materials, Chemical, Value)


## I missed this one !!

sberry %<>% mutate(Chemical = str_trim(paste(Chem_family, Chemical)))

sberry %<>% select(Year, State, Type, Production, Avg, Measures, Materials, Chemical, Value)

### let's look at chemicals being applied to food

unfood <- sberry %<>% filter(Production=="APPLICATIONS") 


unfood %<>% filter(Value != "(D)")

unfood %<>% filter(Value !=  "(NA)")

unfood %<>% filter(Measures == "MEASURED IN LB / ACRE / APPLICATION")

unfood$type %>% unique()                  ## BEARING
unfood$production %>% unique()            ## APPLICATIONS
unfood$Avg %>% unique()                   ## AVG
unfood$Measures %>% unique()              ## MEASURED IN LB / ACRE / APPLICATION
unfood$Materials %>% unique()      ## long list!!        
unfood$Chemical %>% unique()


unfood_1 <- unfood %>%  select(Year, State, Chemical, Value)

unfood_1 %<>% pivot_wider(names_from = Chemical, values_from = Value)

a <- unfood_1$FUNGICIDE


kable(head(sberry, n=10)) %>% kable_styling(font_size=12)

```

## 4.Exploratory data analysis
```{r}
#remove the unuseful value
sberry$Chemical <- ifelse(sberry$Chemical=="", NA, sberry$Chemical)
sberry[is.na(sberry)] <- ""
sberry$Value <- ifelse(sberry$Value==" (D)", NA, sberry$Value)
sberry[is.na(sberry)] <- ""
sberry$Value <- ifelse(sberry$Value==" (NA)", NA, sberry$Value)
sberry[is.na(sberry)] <- ""
sberry$Value <- ifelse(sberry$Value==" (Z)", NA, sberry$Value)
sberry[is.na(sberry)] <- ""
```

## (1)chemical and value
```{r}
#Firstly, I am interested in finding the relationship between the chemical type and value
sberry1<-sberry%>%select(Year,State,Production,Chemical,Value)
sberry1$Value %<>% as.numeric() 

#draw the boxplot to see the distribution of value by using different type of chemical
ggplot(sberry1) +
  geom_boxplot(mapping = aes(x = Chemical, y = Value)) +
  theme(axis.title = element_text(size = 12, face = "bold")) +
  labs(x = "Chemical Type")
```
```{r}
#The plot shows that the range of different chemical is big, and every chemical type has extreme different value here. And also there are many outliers that make the plot not useful, so according to the plot above, I make some adjustment and create another plot.
sberry2 <- filter(sberry1,Chemical=="(NITROGEN)"|Chemical=="(PHOSPHATE)"|Chemical=="(POTASH)")
ggplot(sberry2) +
  geom_boxplot(mapping = aes(x = Chemical, y = Value)) +
  coord_cartesian( ylim=c(0,50))+
  theme(axis.title = element_text(size = 12, face = "bold")) +
  labs(x = "Chemical Type")

sberry3 <- filter(sberry1,Chemical=="FUNGICIDE"|Chemical=="HERBICIDE"|Chemical=="INSECTICIDE")
ggplot(sberry3) +
  geom_boxplot(mapping = aes(x = Chemical, y = Value)) +
  coord_cartesian( ylim=c(0,15))+
  theme(axis.title = element_text(size = 12, face = "bold")) +
  labs(x = "Chemical Type")
#So, we can see the plot above, the number of outliers are huge. When I delete them, I can say that the values of nitrogen and potash are bigger than other types generally. 
```
## (2)State and value
```{r}
# Secondly, I want to confirm that whether price of production was affects by State or not.
ggplot(sberry1)+ 
  geom_boxplot(aes(x=State,y=Value,fill=State))+ 
  coord_cartesian( ylim=c(0,30))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_wrap(.~Year,scales = "free")
#The data of Oregon is small and the value is high, so when I delete the outliers, the value of oregon is missing. We can see that the mean value of Florida is always higher that California in recent two years. The range of the value is very big in Washington.
```

## (3)Year and value
```{r}
#Thirdly,I want to confirm that whether price of production was affects by year or not.
ggplot(sberry1)+ 
  geom_boxplot(aes(x=Year,y=Value,group=Year,fill=Year))+ 
  coord_cartesian( ylim=c(0,10))
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
# The mean value of every year is almost same.
```

## 5.Conclusion

According the dataset and EDA distribution, I can conclude that the correlations of chemical type, state and years are different. But the simple boxplot does not show more right now. I may need to more time to research deeper. 

During this time, I learned how to clean data, and how to organize data. When I did EDA part, I really want to use pca to predict , but there are some errors that need me to fix later. I will review the class recording to fix them.

## 6.Reference
```{r}
citation("ggplot2")
citation("knitr")
citation("magrittr")
citation("kableExtra")
citation("citation")
```
Class recording 14-18, MA615

https://shiny.rstudio.com/tutorial/



